Hysteresis in human computation:\\ how one task affects another

### 2015-02-18 docs/publications/2014.09_SCIENCE/man.tex
small edits, fix bib

\begin{sciabstract}
Microtask platforms are becoming commonplace tools for performing human
research, producing gold-standard data, and annotating large datasets.
These platforms connect \textit{requesters}
(researchers or companies) with large populations (crowds) of workers, who 
perform clerical jobs, typically taking less than five minutes each.
Recently, they have been studied as a way for untrained workers to perform 
clinical diagnostic functions, where expert human resources are
a bottleneck \cite{Warby2014385}.
A topic of ongoing research concerns designing jobs that elicit high 
quality annotations. Here we identify a feature of nearly all crowdsourcing 
jobs that acts as a strong systemic source of bias. 
Microtask jobs typically consist of a sequence 
of tasks sharing common format 
(e.g., circle galaxies in an image). 
Using image-labeling, a canonical microtask job format, we 
discover that a priming effect, induced by earlier tasks, shifts the 
distribution of future answers by 30-50\%. 
Moreover, prior tasks influence the content that workers focus on, 
as well as the richness and specialization of responses. 
Understanding these \textit{intertask effects} is crucial 
to embracing micratask platforms in critical applications.
While intertask effects can be a source of systematic bias, 
our results suggest that, with appropriate job design, 
it might be possible to leverage them to 
hone worker focus and acuity, helping to elicit reproducible, 
expert-level judgments.
\end{sciabstract}

@@

We investigate the potential for task repetition to induce bias in 
a highly generalizeable microtask setup, in which workers label a 
series of images.  If task-induced priming arises in our setup, 
it should be expected in virtually every microtask and crowdsourced workflow
where workers are called upon to exercise their judgment\footnote{
	As opposed to those tasks where a single clear correct answer exists.
}.
Our results show that a worker's response during a task
does indeed depend strongly on the tasks performed beforehand.   
We call the effects that earlier tasks exert on later ones 
\textit{intertask effects}.  If, as has been suggested \cite{5543192}, 
microtask platforms are to be considered as a new form of computing 
architecture, it will be necessary reconcile the fact that the human computing 
elements 
exhibit \textit{hysteresis}, meaning that microtask workers' outputs depend 
on the \textit{history} of their inputs.

@@

In our first experiment, called \textit{intertask-food-objects},
workers were assigned to either a
\textit{food} or \textit{objects} treatment.  Workers in both treatments
performed a set of five \textit{initial tasks}, which each involved 
providing descriptive labels for an image depicting 
either food or (non-food) objects, depending on the worker's 
treatment (Fig.~\ref{fig:task}A and B).  Following the initial 
tasks, workers performed five \textit{test tasks}, which were identical for 
both treatments.  In the the test tasks, workers labeled images depicting a 
mixture of food and objects (Fig.~\ref{fig:task}C).  The five initial 
and test tasks performed by each worker comprised one HIT\footnote{
	Workers were only allowed to participate in one HIT relating to this study.
}. The images were presented one at a time, and we made no 
distinction between the initial and test tasks.

@@

Having affirmed the \textit{existence} of intertask effects, we wish to
understand the nature and severity of the bias thus induced.
We therefore define the extent of bias, $\theta$, in terms of
the strength of influence that a prior perturbation has on workers' responses.
Since workers' responses are stochastic, we define the extent of bias in 
terms of the change induced in the distribution of responses\footnote{
	This is a general divergence metric between probability 
	distributions known as the \textit{total variation distance.}
}:
\begin{equation}
	\theta = \frac{1}{2}\sum_{x \in X} \left| p_0(x) - p_1(x) \right|,
	\label{eq:theta}
\end{equation}
where $p_i(x)$ is the probability that a worker, subjected to treatment $i$,
provides response $x$, from the set of possible responses $X$.
%A significant result from a $\chi^2$ test only allows us to conclude that 
%$\theta$ is non-zero (see supplementary text).  
Determining $\theta$ from finite samples of responses, using purely 
statistical approaches, is difficult (see supplementary text),
which leads us to reformulate the
problem in terms of classification: given a worker's 
responses to test tasks, can a machine reliably infer what her prior exposure 
had been?  Intuitively, the greater the effect that prior tasks have on 
subsequent responses, the more accurate the classifier can be.  
Formally (as shown in the supplementary text), the accuracy of such a 
classifier, $\eta$, establishes a lower bound on the bias:
\begin{equation}
	\theta \geq 2\eta - 1.
	\label{l1}
\end{equation}

@@

\begin{figure}
	\centering
	\includegraphics[scale=0.6]{figs/images.pdf}
	\caption{
		Examples of images used in
		initial tasks for the (\textbf{A}) \textit{food} and (\textbf{B}) 
		\textit{objects} treatments of \textit{intertask-food-objects};
		(\textbf{C}) test tasks for \textit{intertask-food-objects} and 
		\textit{frame-food-objects};
		initial tasks for the (\textbf{D}) \textit{food} and (\textbf{E}) 
		\textit{culture} treatments of \textit{intertask-food-culture};
		and (\textbf{F}) test tasks for \textit{intertask-food-culture} and 
		\textit{frame-food-culture}.
		The full set of experimental materials is shown in the 
		supplementary text.
	}

@@

There has been considerable interest in the effects that framing can have
on microtask responses
\cite{Kinnaird2012281,chandler2013breaking,thibodeau2013natural}, so,
as a point of comparison, we describe a similar experiment 
in which we framed the purpose of the tasks.  In the experiment 
\textit{frame-food-objects}, the same test-tasks were used, but rather
than exposing workers to initial tasks, workers were either told that 
the tasks were ``Funded by the laboratory for the visual perception of Food 
and Ingredients'', or ``\ldots of Objects and Tools''.  
Framing did induce changes in the frequencies of 
word usage at significance (as determined by $\chi^2$ test; $p=0.0012$), but 
the \textit{extent} of framing-induced bias was not statistically 
distinguishable from zero ($p =0.37$) (Fig.~\ref{fig:theta}A), and was
weaker than the bias due to intertask effects ($p=2.1\times 10^{-5}$).

@@

\begin{figure}
	\centering
	\includegraphics[scale=1]{figs/theta.pdf}
	\caption{
		Empirical bias, $\theta_\mathrm{NB}$, measured using a na\"ive Bayes 
		classifier, induced in image labeling tasks, due to workers' 
		exposure to framing and initial tasks.  
		(\textbf{A}) Exposing workers to initial tasks induces more severe
		bias than framing, except when echoed framing is used. 
		(\textbf{B}) As workers proceed through test tasks, 
		the bias due to initial tasks wanes, 
		but remains significant even after five tasks.  
		Standard error bars are shown.
	}
	\label{fig:theta}
\end{figure}

@@

\begin{figure}
	\includegraphics[scale=0.8]{figs/tasks.pdf}
	\caption{Examples of (A) task instructions, (B) a framing slide
		(C) an echoed framing slide, and (D) an image-labeling 
		task.
	}
	\label{fig:hit_preamble}
\end{figure}
Workers were randomly assigned to the treatments, and 
a given worker could only complete one HIT related to this study.  
In \textit{intertask-food-culture}, the \textit{food} and \textit{culture}
treatments were both divided into five sub-treatments, 
in which the test tasks were permuted.  
The permutations were obtained by rotating 
the order shown in Fig.~S\ref{fig:task1:test} by different amounts. 
For example, rotating by one position involves moving the first image to the 
second position, the second image to the third position, and so on, and 
moving the last image to the first position.  
The five different possible rotations were used for the five 
sub-treatments.  This enabled us to study the effects on each image when
it occurred in different positions relative to the initial tasks.

@@

\begin{figure}
	\begin{center}
	\includegraphics{figs/task1-food.pdf}
	\end{center}
	\caption{
		Images used in the initial tasks for the
		\textit{food} treatment of \textit{intertask-food-objects}.  
		The numbers show the order in which the 
		images were presented to workers.
	}
	\label{fig:task1:food}
\end{figure}

\begin{figure}
	\begin{center}
	\includegraphics{figs/task1-obj.pdf}
	\end{center}
	\caption{
		Images used in the initial tasks for the
		\textit{objects} treatment of \textit{intertask-food-objects}.  
		The numbers show the order in which the images were presented to 
		workers.
	}
	\label{fig:task1:obj}
\end{figure}

\begin{figure}
	\begin{center}
	\includegraphics{figs/task1-test.pdf}
	\end{center}
	\caption{
		Images used in the test tasks for \textit{intertask-food-objects}
		and \textit{frame-food-objects}.  
		The numbers show the order in which the 
		images were presented to workers in \textit{frame-food-objects};
		in \textit{intertask-food-objects}, five different orderings were
		used, which can be obtained by taking the numbers shown, $n$,
		and replacing them by $n + c \bmod 5$ for $c$ running from 0 to 4.
	}
	\label{fig:task1:test}
\end{figure}

\begin{figure}
	\begin{center}
	\includegraphics{figs/task2-food.pdf}
	\end{center}
	\caption{
		Images used in the initial tasks for the
		\textit{food} treatment of \textit{intertask-food-culture}.  
		The numbers show the order in which the 
		images were presented to workers.
	}
	\label{fig:task2:food}
\end{figure}

\begin{figure}
	\begin{center}
	\includegraphics{figs/task2-cult.pdf}
	\end{center}
	\caption{
		Images used in the initial tasks for the
		\textit{culture} treatment of \textit{intertask-food-culture}.  
		The numbers show the order in which the 
		images were presented to workers.
	}
	\label{fig:task2:cult}
\end{figure}

\begin{figure}
	\begin{center}
	\includegraphics{figs/task2-test.pdf}
	\end{center}
	\caption{
		Images used in the test tasks for \textit{intertask-food-culture} 
		and \textit{frame-food-culture}.  
		The numbers show the order in which the 
		images were presented to workers.
	}
	\label{fig:task2:test}
\end{figure}

\begin{figure}
	\begin{center}
		\includegraphics{figs/task2-ingr.pdf}
	\end{center}
	\caption{
		Images used in initial tasks for \textit{frame-food-culture} 
		(both for the \textit{food} and \textit{culture} treatments).
		The numbers show the order in which the images were presented to 
		workers.
	}
	\label{fig:frame2:ingr}
\end{figure}

>>>>

\begin{sciabstract}
Microtask platforms are becoming commonplace tools for performing human
research, producing gold-standard data, and annotating large datasets.
These platforms connect \textit{requesters}
(researchers or companies) with large populations (crowds) of workers, who 
perform small jobs, typically taking less than five minutes each.
A topic of ongoing research concerns designing jobs that elicit high 
quality annotations. Here we identify a feature of nearly all crowdsourcing 
jobs that acts as a strong systemic source of bias. 
Microtask jobs typically consist of a sequence 
of tasks sharing common format 
(e.g., circle galaxies in an image). 
Using image-labeling, a canonical microtask job format, we 
discover that earlier tasks have a priming effect on the worker, shifting the 
distribution of future answers by 30-50\%. 
Moreover, prior tasks influence the content that workers focus on, 
as well as the richness and specialization of responses. 
Understanding these \textit{intertask effects} is crucial 
to embracing micratask platforms in critical applications.
While intertask effects can be a source of systematic bias, 
our results suggest that, with appropriate job design, 
they might be leveraged to hone worker focus and acuity, 
helping to elicit reproducible, expert-level judgments.
\end{sciabstract}

@@

We investigate the potential for task repetition to induce bias in 
a highly generalizeable microtask setup, in which workers label a 
series of images.  If task-induced priming arises in our setup, 
it should be expected in virtually every microtask and crowdsourced workflow
where workers are called upon to exercise their judgment\footnote{
	As opposed to those tasks where a single clear correct answer exists.
}.
Our results show that a worker's response during a task
does indeed depend strongly on the tasks performed beforehand.   
We call the effects that earlier tasks exert on later ones 
\textit{intertask effects}.  If, as has been suggested \cite{5543192}, 
microtask platforms are to be considered as a new form of computing 
architecture, it will be necessary to reconcile the fact that the human 
computing elements 
exhibit \textit{hysteresis}, meaning that microtask workers' outputs depend 
on the \textit{history} of their inputs.

@@

In our first experiment, called \textit{intertask-food-objects},
workers were assigned to either a
\textit{food} or \textit{objects} treatment.  Workers in both treatments
performed a set of five \textit{initial tasks}, which each involved 
providing descriptive labels for an image depicting 
either food or (non-food) objects, depending on the worker's 
treatment (Fig.~\ref{fig:task}A and B).  Following the initial 
tasks, workers performed five \textit{test tasks}, which were identical for 
both treatments.  In the the test tasks, workers labeled images, each of 
which depicted a mixture of food and objects (Fig.~\ref{fig:task}C).  
The five initial 
and test tasks performed by each worker comprised one HIT\footnote{
	Workers were only allowed to participate in one HIT relating to this study.
}. The images were presented one at a time, and we made no 
distinction between the initial and test tasks.

@@

Having affirmed the \textit{existence} of intertask effects, we wish to
understand the nature and severity of the bias induced.
We therefore define the extent of bias, $\theta$, in terms of
the strength of influence that a prior perturbation has on workers' responses.
Since workers' responses are stochastic, we define the extent of bias in 
terms of the change induced in the distribution of responses\footnote{
	This is a general divergence metric between probability 
	distributions known as the \textit{total variation distance.}
}:
\begin{equation}
	\theta = \frac{1}{2}\sum_{x \in X} \left| p_0(x) - p_1(x) \right|,
	\label{eq:theta}
\end{equation}
where $p_i(x)$ is the probability that a worker, subjected to treatment $i$,
provides response $x$, from the set of possible responses $X$.
%A significant result from a $\chi^2$ test only allows us to conclude that 
%$\theta$ is non-zero (see supplementary text).  
Determining $\theta$ from finite samples of responses, using purely 
statistical approaches, is difficult (see supplementary text),
which leads us to reformulate the
problem in terms of classification: given a worker's 
responses to test tasks, can a machine reliably infer what her prior exposure 
had been?  Intuitively, the greater the effect that prior tasks have on 
subsequent responses, the more accurate the classifier can be.  
Formally (as shown in the supplementary text), the accuracy of such a 
classifier, $\eta$, establishes a lower bound on the bias:
\begin{equation}
	\theta \geq 2\eta - 1.
	\label{l1}
\end{equation}

@@

\begin{figure}
	\centering
	\includegraphics[scale=1.0]{figs/images.jpg}
	\caption{
		Examples of images used in
		initial tasks for the (\textbf{A}) \textit{food} and (\textbf{B}) 
		\textit{objects} treatments of \textit{intertask-food-objects};
		(\textbf{C}) test tasks for \textit{intertask-food-objects} and 
		\textit{frame-food-objects};
		initial tasks for the (\textbf{D}) \textit{food} and (\textbf{E}) 
		\textit{culture} treatments of \textit{intertask-food-culture};
		and (\textbf{F}) test tasks for \textit{intertask-food-culture} and 
		\textit{frame-food-culture}.
		The full set of experimental materials is shown in the 
		supplementary text.
	}

@@

There has been considerable interest in the effects that framing can have
on microtask responses
\cite{Kinnaird2012281,chandler2013breaking,thibodeau2013natural}, so,
as a point of comparison, we describe a similar experiment 
in which we framed the purpose of the tasks.  In the experiment 
\textit{frame-food-objects}, the same test-tasks were used, but rather
than exposing workers to initial tasks, workers were either told that 
the tasks were ``Funded by the laboratory for the visual perception of Food 
and Ingredients'', or ``\ldots of Objects and Tools''.  
Framing did induce changes in the frequencies of 
word usage at significance (as determined by $\chi^2$ test; $p=0.0012$), but 
the \textit{extent} of framing-induced bias was not statistically 
distinguishable from zero ($p =0.37$) (Fig.~\ref{fig:theta}A), and was
much weaker than the bias due to intertask effects ($p=2.1\times 10^{-5}$).
\begin{figure}
	\centering
	\includegraphics[scale=1]{figs/theta.pdf}
	\caption{
		Empirical bias, $\theta_\mathrm{NB}$, measured using a na\"ive Bayes 
		classifier, induced in image labeling tasks, due to workers' 
		exposure to framing and initial tasks.  
		(\textbf{A}) Exposing workers to initial tasks induces more severe
		bias than framing, except when echoed framing is used. 
		(\textbf{B}) As workers proceed through test tasks, 
		the bias due to initial tasks wanes, 
		but remains significant even after five tasks.  
		Standard error bars are shown.
	}
	\label{fig:theta}
\end{figure}

@@

@@

\begin{figure}
	\includegraphics[scale=0.8]{figs/tasks.jpg}
	\caption{Examples of (A) task instructions, (B) a framing slide
		(C) an echoed framing slide, and (D) an image-labeling 
		task.
	}
	\label{fig:hit_preamble}
\end{figure}
Workers were randomly assigned to the treatments, and 
a given worker could only complete one HIT related to this study.  
In \textit{intertask-food-culture}, the \textit{food} and \textit{culture}
treatments were both divided into five sub-treatments, 
in which the test tasks were permuted.  
The permutations were obtained by rotating 
the order shown in Fig.~S\ref{fig:task1:test} by different amounts. 
For example, rotating by one position involves moving the first image to the 
second position, the second image to the third position, and so on, and 
moving the last image to the first position.  
The five different possible rotations were used for the five 
sub-treatments.  This enabled us to study the effects on each image when
it occurred in different positions relative to the initial tasks.

@@

\begin{figure}
	\begin{center}
	\includegraphics{figs/task1-food.jpg}
	\end{center}
	\caption{
		Images used in the initial tasks for the
		\textit{food} treatment of \textit{intertask-food-objects}.  
		The numbers show the order in which the 
		images were presented to workers.
	}
	\label{fig:task1:food}
\end{figure}

\begin{figure}
	\begin{center}
	\includegraphics{figs/task1-obj.jpg}
	\end{center}
	\caption{
		Images used in the initial tasks for the
		\textit{objects} treatment of \textit{intertask-food-objects}.  
		The numbers show the order in which the images were presented to 
		workers.
	}
	\label{fig:task1:obj}
\end{figure}

\begin{figure}
	\begin{center}
	\includegraphics{figs/task1-test.jpg}
	\end{center}
	\caption{
		Images used in the test tasks for \textit{intertask-food-objects}
		and \textit{frame-food-objects}.  
		The numbers show the order in which the 
		images were presented to workers in \textit{frame-food-objects};
		in \textit{intertask-food-objects}, five different orderings were
		used, which can be obtained by taking the numbers shown, $n$,
		and replacing them by $n + c \bmod 5$ for $c$ running from 0 to 4.
	}
	\label{fig:task1:test}
\end{figure}

\begin{figure}
	\begin{center}
	\includegraphics{figs/task2-food.jpg}
	\end{center}
	\caption{
		Images used in the initial tasks for the
		\textit{food} treatment of \textit{intertask-food-culture}.  
		The numbers show the order in which the 
		images were presented to workers.
	}
	\label{fig:task2:food}
\end{figure}

\begin{figure}
	\begin{center}
	\includegraphics{figs/task2-cult.jpg}
	\end{center}
	\caption{
		Images used in the initial tasks for the
		\textit{culture} treatment of \textit{intertask-food-culture}.  
		The numbers show the order in which the 
		images were presented to workers.
	}
	\label{fig:task2:cult}
\end{figure}

\begin{figure}
	\begin{center}
	\includegraphics{figs/task2-test.jpg}
	\end{center}
	\caption{
		Images used in the test tasks for \textit{intertask-food-culture} 
		and \textit{frame-food-culture}.  
		The numbers show the order in which the 
		images were presented to workers.
	}
	\label{fig:task2:test}
\end{figure}

\begin{figure}
	\begin{center}
		\includegraphics{figs/task2-ingr.jpg}
	\end{center}
	\caption{
		Images used in initial tasks for \textit{frame-food-culture} 
		(both for the \textit{food} and \textit{culture} treatments).
		The numbers show the order in which the images were presented to 
		workers.
	}
	\label{fig:frame2:ingr}
\end{figure}
